\large \textbf{Leveraging Historical Ties between Cognitive Science and Computer Science to Guide Programming Education}

### 2016-02-03 paper.tex

% sorting=none will order citations in order of appearance
% use sorting=nyt to sort by name, year, title
\usepackage[backend=biber,autolang=hyphen, bibencoding=inputenc, isbn=false, uniquename=false, sorting=none, sortcites=true, style=numeric, autocite=superscript, url=false]{biblatex} % biblatex setup
\DeclareLanguageMapping{american}{american-apa}
\appto{\bibfont}{\footnotesize} % for 10pt bibliography
%\usepackage{xunicode}
\usepackage[american]{babel}
\usepackage[babel]{csquotes}
\usepackage[utf8]{inputenc}
\usepackage{float}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage[dvipsnames,table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows,shapes,fit}
\usepackage{standalone}
%\usepackage{mdframed}
\usepackage{changepage}
\usepackage{titlesec}
\usepackage[parfill]{parskip}
\usepackage{mdframed}

@@

%%\let\textcite=\autocite
\addbibresource{computing.bib}

@@

\subsection{Problem Solving}
Developing problem solving was widely regarded as a common objective
for introductory programming courses. Achieving this objective is more
likely if the course begins with a focus on problem solving
strategies, before any programming is
introduced\autocite{koulouri_teaching_2014}. Separating
problem-solving instruction in this way has a number of benefits
grounded in cognitive science research: first, it serves to activate
prior knowledge\autocite{ambrose_how_2010}. With problem-solving
strategies fresh in their memory, learners will more easily apply them
when they begin programming. Second, it helps manage cognitive
load. Where as in the previous example of language choice we reduced
cognitive load by choosing a language with a simpler syntax, here we
separate two tasks that we know individually demand a high level of
cognitive load for novices: learning problem-solving strategies, and
learning programming techniques.

@@

\section{Considerations and Limitations}
In the above hypothetical example we had the luxury of complete
freedom in designing all aspects of an introductory programming
curriculum. We recognize this is almost always never the case in
practice. However, we argue that following a framework such as
constructive alignment as close as possible will help. Thinking about
the framework will also encourage critical engagement with the
constraints imposed by the institution. For example, there has been
much interest in assessment tools used for introductory programming
courses\autocite{lister_blooming_2000,lister_objectives_2001,sheard_exploring_2011},
and in particular automated
assessment\autocite{cheang_automated_2003,traynor_synthesis_2005,ihantola_review_2010}. In
many cases the push towards automated assessment techniques is a
response to increasing class sizes that instructors typically have
little control over. In the event that a large class size dictates
that some form of automated assessment be used, following the
constructive alignment framework would constrain the types of course
objectives and instruction that were feasible. For instance, if the
primary course objective was to teach conceptual understanding of
computer science, then a well-designed multiple-choice assessment
could be used
effectively\autocite{lister_blooming_2000,lister_objectives_2001}.

>>>>

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[babel]{csquotes}
\usepackage{graphicx}
\usepackage{etoolbox}
\usepackage{changepage}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage[parfill]{parskip}
\usepackage[margin=1in]{geometry}

%\usepackage{xunicode}
\usepackage[dvipsnames,table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows,shapes,fit}
\usepackage{standalone}

% sorting=none will order citations in order of appearance
% use sorting=nyt to sort by name, year, title
\usepackage[backend=biber,autolang=hyphen, bibencoding=inputenc, isbn=false, uniquename=false, sorting=none, sortcites=true, style=numeric, autocite=superscript, url=false]{biblatex} % biblatex setup
\addbibresource{./computing.bib}
\appto{\bibfont}{\footnotesize} % for 10pt bibliography

@@

\let\textcite=\autocite

@@

\subsection{Problem solving}
Developing problem solving was widely regarded as a common objective
for introductory programming courses. Achieving this objective is more
likely if the course begins with a focus on problem solving
strategies, before any programming is
introduced\autocite{koulouri_teaching_2014}. Separating
problem-solving instruction in this way has a number of benefits
grounded in cognitive science research: first, it serves to activate
prior knowledge\autocite{ambrose_how_2010}. With problem-solving
strategies fresh in their memory, learners will more easily apply them
when they begin programming. Second, it helps manage cognitive
load. Where as in the previous example of language choice we reduced
cognitive load by choosing a language with a simpler syntax, here we
separate two tasks that we know individually demand a high level of
cognitive load for novices: learning problem-solving strategies, and
learning programming techniques.

@@

\section{Considerations and limitations}
In the above hypothetical example we had the luxury of complete
freedom in designing all aspects of an introductory programming
curriculum. We recognize this is almost always never the case in
practice. However, we argue that following a framework such as
constructive alignment as close as possible will help. Thinking about
the framework will also encourage critical engagement with the
constraints imposed by the institution. For example, there has been
much interest in assessment tools used for introductory programming
courses\autocite{lister_blooming_2000,lister_objectives_2001,sheard_exploring_2011},
and in particular automated
assessment\autocite{cheang_automated_2003,traynor_synthesis_2005,ihantola_review_2010}. In
many cases the push towards automated assessment techniques is a
response to increasing class sizes that instructors typically have
little control over. In the event that a large class size dictates
that some form of automated assessment be used, following the
constructive alignment framework would constrain the types of course
objectives and instruction that were feasible. For instance, if the
primary course objective was to teach conceptual understanding of
computer science, then a well-designed multiple-choice assessment
could be used
effectively\autocite{lister_blooming_2000,lister_objectives_2001}.
